### 2018-11-19 referee_response_v2.txt
first update of referee responses and small text updates to the draft

We would like to thank the referee for the report which helps improve the manuscript, and we have revised the paper accordingly. Detailed replies can be found below.

First of all, we would like to clarify that, the main point of our work is a systematic and comprehensive study of jet quenching in heavy ion collisions using quark (q) and gluon (g) jet substructure as independent probes, instead of what has been summarized in the beginning of the referee report as discrimination between q vs g jets which is only a part of our work. We would also like to clarify that, the study of discriminating quenched and unquenched jets is in fact very natural and at the heart of all jet quenching studies. The goal is to identify and quantitatively describe various aspects of jet modifications, including the dependence on the parton flavor that can lead us towards determining the transport properties of the QGP medium produced in heavy ion collisions. Looking at the q-g classification performance plots is one way to quantify the differences regarding medium modifications to jet (sub)structure which can be captured by the different basis (physics motivated variables, jet images or the telescoping deconstructed variables) one chooses to study jet quenching.

On the other hand, telescoping deconstruction (TD) is developed and applied in this work. In comparison to the original telescoping jets method, that one of us developed a few years ago, this paper presents a crucial improvement leading to TD which is an entirely new framework consisting of jet substructure basis that can in turn be used to represent the various as aspects of jet information. Furthermore, TD is a complete and systematic expansion of jet information utilizing subjets which is analogous to the expansion in perturbation theory, where in the expansion converges rather quickly. We would like to emphasize the following: TD concretely defines all the relevant observables within the framework and gives well-organized jet information. Furthermore, we show in the draft that they capture physical aspects of not only QCD jets but also W and top jets. The role that TD plays in our paper is to provide an answer to how one bridges the conventional approach of studying jets using a list of physics-motivated variables, to modern machine-learning approach of studying jets using neural network to exhaustively extract useful features. This implies that TD should be a very useful and reliable framework for jet physics studies, which on its own is a novel jet representation method. Nevertheless, we describe in detail, the TD method in the Appendix in order not to distract the readers from the main, physical point of our paper regarding flavor dependence of jet quenching in heavy ion collisions.

To answer the referee's questions,

In the original telescoping jets paper, only a single variable called volatility was used to capture certain aspects of jets. On the other hand, TD is a jet substructure basis which is designed to represent, both qualitatively and quantitatively all aspects of jets. As emphasized in the above paragraph, TD is a fixed-order subjet expansion which converges quickly, as implied by the moderate, but characteristic increase of the quark gluon discrimination performance as the order increases. In the case of top jet tagging, TD not only captures the three-prong, hard jet substructure but also other jet features. For example, the color-singlet isolation of W jets and the partonic origin of subjets, which become manifest at N = 2 and N = 3 orders, but are not seen at the N = 1 order.

Figure 18 shows the comparisons of different orders of TD w/ a CNN for the various classification problems such as Top vs QCD, W vs QCD and quark vs gluon jets. If we look at the right most panel, at a given top efficiency of 30% (which is a nominal working point in experiments), we find that the mistag rate ~ 1 in 200 for T1 and as we increase the orders, the rate becomes quite small ~ 1 in 450 which is greater than a 200% improvement. Thus, in going to higher orders of TD, the structure of the decaying top is better represented with encapsulating more of the interesting physics that separates it from a QCD background jet.

A neural network consisting of a Multi-Layer Perceptron with hidden layer(s) (which we describe in the paper) is conveniently used to combine the TD variables and is much simpler than the convolutional neural network that is employed for the jet image method (which we also describe in the paper). The training and testing of the simple neural network can be done in the order O(10s)-O(100s) on most modern laptops. One could have used other MLPs and their variants such as boosted decision tree (BDT) or Random Forests (RF) which are getting quite popular in the literature but that would shift the main focus of the paper to a study of different methods which is not what we intended. The use of multivariate methods available on the market is to primarily combine variables in a non-intuitive linear/non-linear combination (due to the use of non-linear activation functions and hidden layers) and make optimal use of the correlations among them that a traditional analysis using likelihood ratio methods is likely to ignore.

We thank the referee for the raising the question regarding the impact of a realistic underlying event/large background fluctuations. We are aware and intend to perform a further study to address the concern by comparing the q-g discrimination performance for a variety of underlying event descriptions and jet quenching models in a new paper. However, we would like to first clarify that, in our current JEWEL study, the pp and AA jets are represented as a collection of ideally background-subtracted jets. The pp jets are essentially generated by a version of PYTHIA-6 which is implemented in JEWEL simulations without pileup (but with a model of the pp underlying event). The AA jets in JEWEL include both with and without medium recoil contributions and the jets with recoils are background-subtracted using the GridSub method introduced by one of the authors in an earlier publication (in JHEP). However, JEWEL does not generate/simulate the entire heavy ion background, as we have emphasized in the draft, but it includes interactions of the propagating parton with the thermalized medium model implemented in JEWEL. With the application of the GridSub method, one can study an idealized situation of quenched jets which are comparable to unquenched vacuum jets. With the implementation of these procedures, JEWEL describes the jet suppression and a variety of jet substructure modifications (jet shape and splitting function, for example) quite well. In order to extract any meaningful physics statements about flavor dependent jet quenching, it is essential for our model to provide a satisfactory description of data and JEWEL indeed does that. Our envisioned goal for this paper is to present the first quantification of the baseline comparison of quenched quark vs quenched gluon jets and follow up with a study including the underlying event specifically utilizing new MC toolkits that have surfaced in the community since the time of our submission (such as JETSCAPE).

We have mentioned in the paper that the reason why we choose JEWEL for our study is precisely due to its qualitative description of jet quenching measurements at the LHC especially with the inclusions of recoils and the background subtraction methods.

It is indeed correct that pTD and particle multiplicity are IRC unsafe at the particle level. However, the jet constituents used in our work to evaluate all jet observables are essentially grouped into “calorimeter-like towers” with pixel size 0.08 x 0.08 to reflect experimental limitations. This regulates all the IRC singularities and makes our observables IRC safe. Nevertheless, this also implies that jet features that are inherently beyond the angular resolution of 0.08 x 0.08 would be lost in our analysis. This is one of the reasons why the q-g discrimination performances are largely similar amongst the three methods used in this work. Furthermore, these observables such as pTD, have indeed been measured in heavy ion collisions by the ALICE collaboration and JEWEL does a good job of reproducing the structure as seen in data without any additional tunings whatsoever. There have also been publications by both CMS and ATLAS with the use of multiplicity and pTD specifically in the use of q/g classifications.

We would like to clarify the comment from the referee about the use of soft drop observables. We used TD to illuminate the inner working of the soft drop algorithm by comparisons between two methods. Taking into account the limitations imposed by soft drop, and we believe that TD is more flexible and robust by having an additional handle in the order at which TD is calculated. We find that TD is even more related to a variation of soft drop called recursive soft drop, and one of us had many discussions with the recursive soft drop authors. We would also like to clarify that, soft drop is not just a grooming method, and TD can be used as a grooming method. Soft Drop has been used to define observables which are physically meaningful, for example, the groomed momentum sharing zg or the splitting function. On the other hand, TD defines exclusively the subjet regions and disregards radiation in the outside region, which is exactly a grooming procedure.

In fact, what soft drop does is similar to TD by identifying the two subjet branches while dropping soft, wide angle radiation. Since soft drop is based on the C/A de-clustering tree which is angular ordered, the size of the subjet branches has to be smaller than the angle between the branches. This is the reason why in the comparison of TD to soft drop we choose the subjet size to be proportional to the angle between the subjet axes. The prefactor of 0.3, which can surely be varied, offers an additional degree of freedom within the TD framework to scan the energy distribution surrounding the individual subjets. We do vary the prefactor in our studies and show the case with the prefactor 0.3 just as an example, and it is not meant to be a fixed number.

We would like to emphasize again that, the main point of our paper is to use quark and gluon jet substructure to probe heavy ion collisions, instead of simply performing q-g discrimination. It is then very natural to discuss the physical meaning of the TD variables which we use to study jet quenching. The same clarification applies also to the use of the Lund diagram technique, which is another jet substructure representation closely related to the C/A re-clustered tree and soft drop. Our use of Lund diagrams in this paper is to drive home to point of differences between q/g vacuum/quenched jets. Since the Lund diagram is necessarily crafted by using soft drop recursively on the leading prong, it records the leading branch and its kinematics is intimately related to TD, which represents jets using subjet. In the draft, we have highlighted this specific point and tried to make clear that the physics picture we learn from TD and physics variables are also reflected in the Lund diagram. We have modified the draft to better reflect the discussion to the reader.

We show how jet modifications can be probed in the TD framework by looking at the jet substructure observables order-by-order, and at each order we see new information. We show that the information zg carries in soft drop can be studied in the TD framework more generally, and we further look at the angular distribution of subjets, as well as subjet substructure like the subjet masses. Even more, we can construct higher-level observables using the basis observables and give an example of delta m^2 between groomed and ungroomed jet masses, and these observable shows striking signatures of jet quenching as modelled in JEWEL by highlighting the differences between quark and gluon jets. Such an observable could only be created by looking at the impact of q-g discrimination on TD at various orders. We have updated the draft to clearly describe the usefulness of these comparisons and that we urge different experiments to measure these observables to make definitive statements on quark and gluon energy loss.

With reference to the referee’s comments that it is not clear that TD systematically improves q/g discrimination leading to insights about jet quenching mechanism as implemented in MC simulations, we note the following. Figure 13 and 14 shows the comparisons of the ROC curves for the various methods used in the analysis and we find that the TD at 3rd order performs almost as good as the DCNN if not better. With the use of TD, we can extract the corresponding variables that facilitate this performance and these extractions are done in the fragmentation basis of the jet structure (as we discussed in the above paragraphs and in the paper). Since it is quite complicated to extract any physics from the convolution layers in a DCNN, TD provides an excellent option to study inherent differences without any biases.

One of the main reasons why we showed performance plots for discriminating between quenched and unquenched jets is to study the effect of quenching between q vs g jets as implemented in JEWEL. As we mentioned in the paper and shown in Figure 17, all the methods showcase better performance for differentiating quenched gluon jet from an unquenched gluon jet as opposed to quark jets, leading to our conclusion that gluon jets are further modified. We have edited the text to better reflect on a use case for this particular study.

It is a qualitative feature of JEWEL-simulated heavy ion collisions that the q-g discrimination performance decreases from pp to AA. We are actively invested in studying how other jet quenching models and calculations reflect in our methods. Such a study offers us to the opportunity to solve the inverse problem of mapping from jet modifications to the inner working of jet quenching using a consistent framework that we have introduced in this paper.

Sincerely,